---
title: "R Notebook"
output: html_notebook
---

Diamond dataset is inside the ggplot library.

```{r}
library(ggplot2)
data=diamonds
```

```{r}
#Getting structure of Diamond dataset
str(data)
```
 Cut, Color and Clarity are factor variables and other are numerical variables.
 
```{r}
# Histogram of price 
ggplot(aes(x=price),data=diamonds)+geom_histogram()
```
 Histogram is skewed right skewed.
 
```{r}
#Summary statistics 
summary(diamonds$price)
```
 
 Answering the following questions 
1. How many cost less than U$500?
2. How many cost less than U$250?
3. How many cost equal to U$15,000 or more?

```{r}
#Cost less than US$500
sum(diamonds$price<500)
```
```{r}
#Cost less than US$250
sum(diamonds$price<250)
```
```{r}
#cost equal to U$15,000 or more
sum(diamonds$price>=15000)
```
```{r}
# Explore the largest peak in the
ggplot(aes(x=price),data=diamonds)+geom_histogram(binwidth = 1000,col='red')+ggtitle('Histogram of the price')+ylab('Frequency')+xlab('Diamond price')
```
```{r}
# Break out the histogram of diamond prices by cut.
ggplot(aes(x=price),data=diamonds)+geom_histogram(binwidth = 500,col='red')+ggtitle('Histogram of the price')+ylab('Frequency')+xlab('Diamond price')+facet_wrap(~cut)+theme_minimal()
```
```{r}
#Higest price diamond
subset(diamonds,price==max(price))
```

Premimum cut has maximum price diamond.

```{r}
#Lowest Price Diamond
subset(diamonds,price==min(price))
```
Ideal and premium has lowest price diamonds.

To find lowest mean of the diamond cuts 
```{r}
#Subsetting the diamonds by cut
Fair = diamonds[which(diamonds$cut == "Fair"),]
Good = diamonds[which(diamonds$cut == "Good"),]
VaryGood = diamonds[which(diamonds$cut == "Very Good"),]
Premium = diamonds[which(diamonds$cut == "Premium"),]
Ideal = diamonds[which(diamonds$cut == "Ideal"),]

```

```{r}
mean(Fair$price)
```
```{r}
mean(Good$price)
```
```{r}
mean(VaryGood$price)
```
```{r}
mean(Premium$price)
```
```{r}
mean(Ideal$price)
```

In the previous histogram, Scles of y was same for all the cuts. So it was hard to interpret from graph. Now we are changing the scale by just adding scales=free_y

```{r}
ggplot(aes(x=price),data=diamonds)+geom_histogram(binwidth = 500,col='red')+ggtitle('Histogram of the price')+ylab('Frequency')+xlab('Diamond price')+facet_wrap(~cut,scales="free_y")+theme_minimal()
```
 Now figure out price per caret by cut.
```{r}
#Histogram of price per caret by cut
ggplot(aes(x=price/carat),data=diamonds)+geom_histogram(binwidth = 500,col='red')+ggtitle('Histogram of the price per carat')+ylab('Frequency')+xlab('Diamond price per carat')+facet_wrap(~cut,scales="free_y")+theme_minimal()
```
 
 Using log10 for x 
```{r}
ggplot(aes(x=price/carat),data=diamonds)+geom_histogram(binwidth = 0.1,col='red')+ggtitle('Histogram of the price per carat')+ylab('Frequency')+xlab('Diamond price per carat')+facet_wrap(~cut,scales="free_y")+theme_minimal()+scale_x_log10()
```
 
 
 
```{r}
#Plot price and carat by cut
ggplot(aes(x=price,y=carat),data=diamonds)+geom_line()+ylab('carat')+xlab('Diamond price')+facet_wrap(~cut,scales="free_y")+theme_minimal()
```
 
 Now it's term of some interesting boxplots 
```{r}
# Investigate the price of diamonds using box plots
ggplot(diamonds,aes(factor(cut),price,fill=cut))+geom_boxplot()+ggtitle('Boxplot of price by cut')
```
```{r}
# Investigate the price of diamonds using box plots
ggplot(diamonds,aes(factor(color),price,fill=color))+geom_boxplot()+ggtitle('Boxplot of price by color')
```
```{r}
#Subsetting the diamonds by color
D = subset(diamonds,diamonds$color == "D")
E = subset(diamonds,diamonds$color == "E")
F = subset(diamonds,diamonds$color == "F")
G = subset(diamonds,diamonds$color == "G")
H = subset(diamonds,diamonds$color == "H")
I = subset(diamonds,diamonds$color == "I")
J = subset(diamonds,diamonds$color == "J")
```
 
```{r}
summary(D)
```
```{r}
summary(J)
```
```{r}
#IQR of best color
IQR(D$price)
```
```{r}
#IQR of worst color
IQR(J$price)
```
```{r}
# Investigate the price per carat of diamonds using box plots
ggplot(diamonds,aes(factor(color),price/carat,fill=color))+geom_boxplot()+ggtitle('Boxplot of price by color')
```

```{r}
#Frequency polygon
ggplot(data=diamonds, aes(x=carat)) + geom_freqpoly() + ggtitle("Diamond Frequency by Carat") 
```

```{r}
# scatterplot of price vs x.
ggplot(data=diamonds,aes(x=price,y=x))+geom_point()
```
```{r}
#Correlation of price and x
cor.test(data$price,data$x)
```

```{r}
#Correlation of price and y
cor.test(data$price,data$y)
```

```{r}
#Correlation of price and Z
cor.test(data$price,data$z)
```

```{r}
#Create a simple scatter plot of price vs depth
ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha=1/100)+scale_x_continuous(breaks=seq(50,80,1))
```
```{r}
#Correlation of depth and price
cor.test(diamonds$depth,diamonds$price)
```

```{r}
#Create a scatterplot of price vs carat
#and omit the top 1% of price and carat
ggplot(aes(carat,price),data=diamonds)+geom_point(position = position_jitter(h=0))
```
```{r}
# Create a scatterplot of price vs. volume (x * y * z)
# Create a new variable for volume in the diamonds data frame.
diamonds$volume=diamonds$x*diamonds$y*diamonds$z
ggplot(data=diamonds,aes(x=volume,y=price))+geom_point()
```
```{r}
#Count of diamonds whoes volume 0 and greater than 800
library(dplyr)
diamond_subset=filter(diamonds,!( diamonds$volume >=800 | diamonds$volume==0 ))
cor.test(diamond_subset$volume,diamond_subset$price)
```
```{r}
#Scatterplot of volume and price excluding volume 0 and greater than 800
ggplot(aes(x=volume,y=price),data=diamond_subset)+geom_point()+geom_smooth()
```
```{r}
#the data frame diamondsByClarity

diamonds_clarity=group_by(diamonds,clarity)
diamondsByClarity=summarise(diamonds_clarity,clarity_maen=mean(as.numeric(clarity)),clarity_median=median(as.numeric(clarity)),n=n())
```
```{r}
#First top 6 rows 
head(diamonds,6)
```
```{r}
#last top 6 rows 
tail(diamondsByClarity,6)
```
```{r}
# Group by clarity and color
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```

```{r}
#Barplot of clarity

clarity=ggplot(aes(x=clarity,y=mean_price),data=diamonds_mp_by_clarity)+geom_bar(stat="identity",col='red')
color=ggplot(aes(x=color,y=mean_price),data=diamonds_mp_by_color)+geom_bar(stat="identity",col='blue')


```

```{r}
#Histogram of price of different colors
ggplot(aes(x=price),data = diamonds)+geom_histogram(binwidth = 500)+facet_wrap(~color)+scale_fill_brewer(type='qual')
```
```{r}
# Create a scatterplot of diamond price vs cut
ggplot(aes(x=price,y=table),data=diamonds)+geom_point()+scale_color_brewer(color,type = 'qual')
```

```{r}
#scatterplot of diamond price vs volumn
ggplot(aes(x=price,y=log(x*y*z),color=clarity),data=diamonds)+geom_point()+scale_color_brewer(type='div')
```
```{r}
#Scatterplot of price and weight
ggplot(data=diamonds,aes(x=carat,y=price))+geom_point(color='blue',alpha=1/4)+scale_x_continuous(lim = c(0,quantile(diamonds$carat,0.99)))+scale_y_continuous(lim = c(0,quantile(diamonds$price,0.99)))+stat_smooth(method = 'lm',color='red')
```

```{r}

library(GGally)
library(scales)
set.seed(123)
diamond_shap=diamonds[sample(1:length(diamonds$price),10000),]
ggpairs(diamond_shap)
```
```{r}
library(gridExtra)

plot1 <- ggplot(aes(x=price),data=diamonds)+geom_histogram(binwidth = 0.5,color='red')

plot2 <- ggplot(aes(x=log(price)),data=diamonds)+geom_histogram(binwidth = 0.05,color='blue')

grid.arrange(plot1,plot2,ncol=2)
```
```{r}
#Tranforming plot 
ggplot(aes(x=carat,y=price),data=diamonds)+geom_point(color='blue')+scale_y_continuous(trans = log10_trans())
```
```{r}
#Carat table
head(sort(table(diamonds$carat),decreasing = T))
```
```{r}
#Price table
head(sort(table(diamonds$price),decreasing = T))
```
```{r}
#Liner model
m1=lm(I(log(price))~ I(carat^(1/3))+carat+cut+clarity,data=diamonds)
summary(m1)

```







